---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# scmsens

<!-- badges: start -->
[![R-CMD-check](https://github.com/sooahnshin/scmsens/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sooahnshin/scmsens/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Overview

`scmsens` provides sensitivity analysis tools for the Synthetic Control Method (SCM) 
framework to test the robustness of SCM results to potential unobserved donor units, 
as discussed in [Liu, Shin, and Yamauchi (2024)](https://sooahnshin.com/SCM_Missing.pdf). 
It extends the traditional omitted variable bias framework developed by 
[Cinelli and Hazlett (2020)](https://doi.org/10.1111/rssb.12348)
for sensitivity analysis.

## Installation

You can install the development version of scmsens from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("sooahnshin/scmsens")
```

## Example 1: Sensitivity Analysis with Leave-One-Out

```{r loo}
library(scmsens)
library(dplyr)
library(ggplot2)
theme_set(theme_bw(base_size = 15) + theme(plot.title = element_text(hjust = 0.5)))

## Synthetic data generated with interactive fixed effect model
synth_data
# synth_pre/synth_post are preperiod/postperiod data from this synthetic data

## Leave-one-out sensitivity analysis
# We estimate bias parameters with dropping one control unit at a time
res <- estimate_params(var_y_name = "Y", # name of the treated unit
                       var_x_name = paste0("X", 1:16), # names of control units
                       data_pre = synth_pre,
                       data_post = synth_post)
knitr::kable(head(res) |> mutate(across(where(is.numeric), ~ round(.x, 3))))

## Contour plot (SCM Estimate)
tau_complete <- res |> pull(tau) |> unique()
var_x_gamma <- res |> pull(gamma)
var_x_delta <- res |> pull(delta)
var_x_tau <- res |> pull(estimate)
var_x_name <- paste0("Drop ", paste0("X", 1:16), " (", round(var_x_tau, 1), ")")
plot_sensitivity(tau_complete = tau_complete,
                 var_x_gamma = var_x_gamma, var_x_delta = var_x_delta, var_x_name = var_x_name,
                 title = "Sensitivity of the SCM Estimate", text_size = 4)

## Contour plot (R squared)
fit <- vertreg_stacked(
  formula = paste("Y ~ ", paste(paste0("X", 1:16), collapse = " + ")),
  data_pre = synth_pre,
  data_post = synth_post
  )
df <- fit$df # degree of freedom
tau_se <- fit$std.error # standard error
# plot only top 5 units for benchmark (positive bias)
res <- res |> arrange(desc(bias)) |> head(5)
tau_complete <- res |> pull(tau) |> unique()
var_x_r2_Y_Z <- res |> pull(r2_Y_Z)
var_x_r2_D_Z <- res |> pull(r2_D_Z)
var_x_name <- paste0("Drop ", paste0("X", 1:5))
plot_sensitivity_r2(tau_complete = tau_complete, df = df, tau_se = tau_se,
                    sign_bias = 1,
                    var_x_r2_Y_Z = var_x_r2_Y_Z, var_x_r2_D_Z = var_x_r2_D_Z, var_x_name = var_x_name,
                    title = "Sensitivity of the SCM Estimate (Positive Bias)",
                    r2_Y_Z_seq = seq(0, 0.1, length.out = 100),
                    r2_D_Z_seq = seq(0, 0.1, length.out = 100),
                    text_size = 5)

```

## Example 2: Sensitivity Analysis with Partially Observed Data
```{r partial}
## Suppose X1 is partially observed (only observed for odd-numbered rows)
synth_pre_missing <- synth_pre |>
  mutate(X1 = ifelse(row_number() %% 2 == 0, NA, X1)) 
## Estimate parameters with partially observed data 
df_pre_partial <- synth_pre_missing |>
  filter(!is.na(X1))
res <- estimate_params_partial(
  fm_z_on_x = paste("X1 ~ ", paste(paste0("X", 2:16), collapse = " + ")),
  fm_y_on_z_and_x = paste("Y ~ ", paste(paste0("X", 1:16), collapse = " + ")),
  data_pre = df_pre_partial,
  data_post = synth_post,
  pseudo_inverse = FALSE
  )
res_table <- tibble(
  Parameter = names(res),
  Estimate = unlist(res)  # Convert list values to numeric vector
)
knitr::kable(res_table, digits = 3)

## Contour plot (SCM Estimate)
tau_complete <- vertreg(
  formula = paste("Y ~ ", paste(paste0("X", 2:16), collapse = " + ")),
  data_pre = synth_pre_missing,
  data_post = synth_post
  )
plot_sensitivity(tau_complete = tau_complete,
                 var_x_gamma = res$gamma, var_x_delta = res$delta, 
                 var_x_name = paste0("X1 (Partially observed; ", round(tau_complete - res$bias, 1), ")"),
                 title = "Sensitivity of the SCM Estimate", text_size = 4)
```

