---
title: "simulation_study"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simulation_study}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

In this vignette, we conduct a simulation study to evaluate bias in different approaches to the synthetic control method with missing outcomes. Our primary goal is to compare the root mean squared error of causal estimates across the following methods:

- **Complete observations**: Vertical regression using all available data.
- **Complete `gsynth`**: Interactive fixed effects (IFE) model estimation by [Xu (2017)](https://doi.org/10.1017/pan.2016.2) using all the available data.
- **Unit-wise deletion**: Excluding units with missing outcomes.
- **Mean imputation**: Imputing missing values with the sample mean of the partially observed outcomes for the excluded unit.
- **Median imputation**: Imputing missing values with the sample median of the partially observed outcomes for the excluded unit.
- **Bias correction**: Applying the proposed bias decomposition to the observed portion of the outcome data.

We generate simulated data using a linear factor model and the West German reunification dataset. For further details, see Appendix B of the [paper](https://sooahnshin.com/SCM_Missing.pdf).

```{r setup}
library(scmsens)
library(gsynth)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(purrr)
library(ggrepel)
library(splines)
theme_set(theme_bw(base_size = 15) + theme(plot.title = element_text(hjust = 0.5)))
```

# Tutorial

A single simulation run is organized as follows:

1. Fit the IFE model using the West German reunification dataset.
1. Generate synthetic data based on the estimated parameters with augmentation.
1. Introduce missing outcomes and apply the specified methods to estimate the causal effect and assess bias.

## 1. Fit the IFE model

We fit the IFE model using the West German reunification dataset.

```{r ife_model}
data("wgermany")
# Create a long format dataset
wgermany_long <- wgermany |>
  pivot_longer(cols = !(year|Dt), names_to = "country", values_to = "gdp") |>
  mutate(treat = if_else(year >= 1990 & country == "West Germany", 1, 0))
# Fit gsynth
set.seed(123)
gsynth_fit <- gsynth(gdp ~ treat,
    data = wgermany_long,
    index = c("country", "year"),
    force = "none", # we don't add additional unit or time fixed effects
    # A cross-validation procedure is provided (when CV = TRUE)
    # to select the number of unobserved factors within the interval of r = c(0,2).
    CV = TRUE, r = c(0, 2),
    se = TRUE, inference = "parametric", nboots = 1000,
    parallel = FALSE
)
# Check the estimated factors
# plot(gsynth_fit, type = "factors", xlab = "Year")

# Save estimated factors and loadings (using potential outcome Y(0))
wgerm_factor <- gsynth_fit$factor
wgerm_loading_co <- gsynth_fit$lambda.co
wgerm_loading_tr <- gsynth_fit$lambda.tr
wgerm_loading <- rbind(wgerm_loading_co, wgerm_loading_tr) |>
  as.data.frame()

# Check: plot loadings
wgerm_loading %>%
    mutate(unit = rownames(.)) |>
    mutate(treat = if_else(unit == "West Germany", "Treated", "Control")) |>
    ggplot(aes(x = r1, y = r2, color = treat)) +
    scale_color_manual(
        name = "Group", labels = c("Control", "Treated"),
        values = c("black", "indianred")
    ) +
    theme(legend.position = "none") +
    geom_point() +
    geom_label_repel(
        aes(label = unit),
        box.padding = 0.35, # Padding around the label
        point.padding = 0.5, # Avoids overlapping of point and label
        segment.color = 'grey50' # Color of the line connecting point and label
    ) +
    labs(title = "Factor Loadings", x = "Factor 1", y = "Factor 2")
```

To generate synthetic data with longer periods, we augment the estimated factors using a spline model and stochastic noise.

```{r augment_factors}
# The estimated factors
factor_df <- data.frame(
    time = 1:nrow(wgerm_factor),
    factor1 = wgerm_factor[, 1],
    factor2 = wgerm_factor[, 2]
)
# Fit spline model for the augmentation
mod_factor1 <- lm(factor1 ~ bs(time, df = 5), data = factor_df)
mod_factor2 <- lm(factor2 ~ bs(time, df = 6), data = factor_df)
# We will be augmenting the factors in each pre-/post-period
pre_time <- -c(rev(seq(0, floor((99 - nrow(wgerm_factor)) * (3 / 4)))))
post_time <- seq(nrow(wgerm_factor) + 1, length = round((99 - nrow(wgerm_factor)) / 4))
# Augment the factors
aug_factor1 <- predict(mod_factor1, newdata = data.frame(time = c(pre_time, post_time)))
aug_factor2 <- predict(mod_factor2, newdata = data.frame(time = c(pre_time, post_time)))
# Add stochastic noise
set.seed(123)
aug_factor1 <- aug_factor1 + rnorm(length(aug_factor1), 0, 100)
aug_factor2 <- aug_factor2 + rnorm(length(aug_factor2), 0, 100)
# Combine the augmented factors with the original estimated factors
aug_factor_df <- rbind(
    factor_df,
    data.frame(
        time = c(pre_time, post_time),
        factor1 = aug_factor1,
        factor2 = aug_factor2
    )
)
aug_factor_df <- arrange(aug_factor_df, time)
# Check: plot factors
aug_factor_df |>
    pivot_longer(-time, names_to = "factors", values_to = "value") |>
    mutate(factors = str_remove(factors, "factor")) |>
    mutate(time = time - min(time) + 1) |>
    ggplot(aes(x = time, y = value, color = factors)) +
    geom_rect(aes(xmin = length(pre_time) + 1, xmax = length(pre_time) + nrow(wgerm_factor), ymin = -Inf, ymax = Inf),
              fill = "grey90", color = NA, alpha = 0.03) +
    geom_line() +
    labs(title = "Factors", x = "Time", y = "Value") +
    scale_color_brewer(NULL, type = "qual", palette = "Set1") +
    theme(legend.position = "none") 
```

## 2. Generate synthetic data

Now, we generate synthetic data based on the estimated parameters with augmentation.
We assume that the treatment effect for West Germany, the treated unit, is zero.

```{r generate_synthetic_data}
# Setup
tau_true <- 0 # treatment effect for West Germany
n_control <- (nrow(wgerm_loading) - 1) # number of control units
t0 <- (nrow(aug_factor_df) - 1) # the length of pre-treatment period
# Generate data using fitted model (using all countries)
set.seed(1)
synth_ife_data <- generate_synth_data_ife(
    t0 = t0,
    n = n_control,
    tau = tau_true,
    phi = wgerm_loading |> as.matrix(),
    mu = rbind(aug_factor_df$factor1, aug_factor_df$factor2)
)
```

## 3. Introduce missing outcomes and apply methods

We introduce missing outcomes for a unit to simulate missing data, and then apply the specified methods to estimate the causal effect and assess bias.
Here, we introduce helper functions as follows.

```{r helper_functions}
#' Run IFE simulation
#' 
#' @param synth_ife A synth_ife object generated by generate_synth_data_ife()
#' @param miss_index A numeric vector of indices of missing time points
#' @param k An integer of the number of missing units
#' @param include_gsynth A logical indicating whether to include gsynth in the simulation
run_ife_simulation <- function(synth_ife, miss_index = 1:50, k = 2, include_gsynth = FALSE) {
    ## check input type
    if (!isTRUE(class(synth_ife) == "synth_ife")) {
        stop("synth_ife should be a synth_ife object generated by generate_synth_data_ife()")
    }

    ## inputs
    tau_true <- synth_ife$tau
    df_pre <- synth_ife$df_pre_full
    t0 <- nrow(df_pre)
    n_control <- ncol(df_pre) - 1
    df_post <- synth_ife$df_prepost_full[(t0 + 1), ]
    donor_units <- paste0("X", 1:n_control)

    if (isTRUE(length(miss_index) > t0)) {
        stop("miss_index should be a numeric vector of length less than t0")
    }

    # fit the vertical regression with entire data
    fit_complete <- vertreg(
        formula = paste("Y ~  ", paste0("X", 1:n_control, collapse = "+")),
        data_pre = df_pre,
        data_post = df_post |> select(-D)
    )

    if (isTRUE(include_gsynth)) {
        # fit gsynth with entire data
        df_pre_long <- df_pre %>%
            mutate(time = 1:nrow(.)) |>
            pivot_longer(cols = -time, names_to = "unit", values_to = "outcome")
        df_prepost_long <- df_post |>
            select(-D) |>
            mutate(time = t0 + 1) |>
            pivot_longer(cols = -time, names_to = "unit", values_to = "outcome") |>
            mutate(treat = if_else(unit == "Y", 1, 0)) |>
            bind_rows(df_pre_long |> mutate(treat = 0))
        complete_gsynth_mod <- gsynth::gsynth(outcome ~ treat,
            data = df_prepost_long,
            index = c("unit", "time"),
            force = "none", # we don't add additional unit or time fixed effects
            # A cross-validation procedure is provided (when CV = TRUE)
            # to select the number of unobserved factors within the interval of r=c(0,5).
            CV = TRUE, r = c(0, 2),
            se = FALSE, inference = "parametric"
        )
        fit_gsynth_complete <- complete_gsynth_mod$att.avg
    } else {
        fit_gsynth_complete <- NA
    }
    
    # generate all the possible missing combinations
    combinations_result <- generate_combinations(donor_units, k)

    ## Warning for large number of combinations
    if (length(combinations_result) > 500) {
        proceed_or_not()
    }

    # compute the vertical regression results
    res_vert <- map(combinations_result, get_vertreg_results,
        ## arguments
        miss_index = miss_index, df_pre = df_pre,
        df_post = df_post,
        n_control = n_control, t0 = t0
    ) %>%
        bind_rows()

    sim_res <- data.frame(
            estimator = c("True Effect", "Complete Observations", "Complete gsynth"),
            tau = c(tau_true, fit_complete, fit_gsynth_complete),
            gamma = c(0, 0, 0),
            delta = c(0, 0, 0),
            bias_est = c(0, 0, 0),
            missing_var = NA,
            stringsAsFactors = FALSE
        ) |>
            bind_rows(res_vert)

    return(sim_res)
}
#' Vertical regression results with missing values
#' 
#' Iterates over the missing variables and computes the vertical regression results with different approaches (e.g., imputation methods, bias correction).
#' 
#' @param miss_var_name A character vector of missing variable names
#' @param miss_index A numeric vector of indices of missing time points
#' @param df_pre A data frame of the pre-treatment period
#' @param df_post A data frame of the post-treatment period
#' @param n_control An integer of the number of control units
#' @param t0 An integer of the length of the pre-treatment period
get_vertreg_results <- function(miss_var_name, miss_index, df_pre, df_post, n_control, t0) {
    single_miss_flag <- isTRUE(length(miss_var_name) == 1)
    walk(miss_var_name, ~ {
        df_pre <<- df_pre %>%
            mutate("{.x}_miss" := if_else(row_number() %in% miss_index, NA_real_, !!sym(.x)))
    })
    walk(miss_var_name, ~ {
        df_post <<- df_post %>%
            mutate("{.x}_miss" := !!sym(.x))
    })

    obs_var_names <- paste0("X", 1:n_control) |> setdiff(miss_var_name)

    ##
    ## Estimate Bias with Partial Observations
    ##
    # Fit removing the missing X_miss
    fit_missing <- vertreg(
        formula = paste("Y ~  ", paste0(obs_var_names, collapse = "+")),
        data_pre = df_pre,
        data_post = df_post |> select(-D)
    )
    if (single_miss_flag) {
        # Estimate bias with complete observations in X_miss (infeasible)
        bias_est_complete <- estimate_params_partial(
            fm_z_on_x = paste(miss_var_name, "~  -1 + ", paste0(obs_var_names, collapse = "+")),
            fm_y_on_z_and_x = paste0("Y ~  -1 + ", paste0(c(obs_var_names, miss_var_name), collapse = "+")),
            data_pre = df_pre,
            data_post = df_post |> select(-D)
        )

        # Estimate bias with missing value in X_miss (feasible)
        bias_est_missing <- estimate_params_partial(
            fm_z_on_x = paste(paste0(miss_var_name, "_miss"), "~ -1 + ", paste0(obs_var_names, collapse = "+")),
            fm_y_on_z_and_x = paste0("Y ~ -1 + ", paste0(miss_var_name, "_miss"), "+", paste0(obs_var_names, collapse = "+")),
            data_pre = df_pre,
            data_post = df_post |> select(-D)
        )
    } else {
        # Estimate bias with complete observations in X_miss (infeasible)
        bias_est_complete <- estimate_params_partial_multi(
            var_z_name = miss_var_name,
            var_x_name = obs_var_names,
            var_y_name = "Y",
            data_pre = df_pre,
            data_post = df_post |> select(-D)
        )

        # Estimate bias with missing value in X_miss (feasible)
        bias_est_missing <- estimate_params_partial_multi(
            var_z_name = paste0(miss_var_name, "_miss"),
            var_x_name = obs_var_names,
            var_y_name = "Y",
            data_pre = df_pre,
            data_post = df_post |> select(-D)
        )
    }
    ##
    ## Simple imputation
    ##

    # Create long format data for imputation
    df_pre_long <- df_pre %>%
        mutate(time = 1:nrow(.)) |>
        pivot_longer(cols = -time, names_to = "unit", values_to = "outcome") |>
        filter(!unit %in% miss_var_name)
    df_prepost_long <- df_post |>
        select(-D) |>
        mutate(time = t0 + 1) |>
        pivot_longer(cols = -time, names_to = "unit", values_to = "outcome") |>
        filter(!unit %in% miss_var_name) |>
        mutate(treat = if_else(unit == "Y", 1, 0)) |>
        bind_rows(df_pre_long |> mutate(treat = 0))

    # Mean imputation
    df_prepost_long_imp <- df_prepost_long |>
        group_by(unit) |>
        mutate(outcome = if_else(is.na(outcome), mean(outcome, na.rm = TRUE), outcome)) |>
        ungroup()
    df_pre_imp <- df_prepost_long_imp |>
        filter(time <= t0) |>
        select(-treat) |>
        pivot_wider(names_from = "unit", values_from = "outcome") |>
        select(-time)
    df_post_imp <- df_prepost_long_imp |>
        filter(time > t0) |>
        select(-treat) |>
        pivot_wider(names_from = "unit", values_from = "outcome") |>
        select(-time)

    # Median imputation
    df_prepost_long_imp_med <- df_prepost_long |>
        group_by(unit) |>
        mutate(outcome = if_else(is.na(outcome), median(outcome, na.rm = TRUE), outcome)) |>
        ungroup()
    df_pre_med <- df_prepost_long_imp_med |>
        filter(time <= t0) |>
        select(-treat) |>
        pivot_wider(names_from = "unit", values_from = "outcome") |>
        select(-time)
    df_post_med <- df_prepost_long_imp_med |>
        filter(time > t0) |>
        select(-treat) |>
        pivot_wider(names_from = "unit", values_from = "outcome") |>
        select(-time)

    fit_imp <- vertreg(
        # e.g. "Y ~ X3 + ... + X16 + X1_miss + X2_miss"
        formula = paste("Y ~  ", paste0(obs_var_names, collapse = "+"), "+", paste0(paste0(miss_var_name, "_miss"), collapse = "+")),
        data_pre = df_pre_imp,
        data_post = df_post_imp |> as.data.frame() # tibble results in duplicated outputs at the end
    )

    fit_imp_med <- vertreg(
        formula = paste("Y ~  ", paste0(obs_var_names, collapse = "+"), "+", paste0(paste0(miss_var_name, "_miss"), collapse = "+")),
        data_pre = df_pre_med,
        data_post = df_post_med |> as.data.frame()
    )


    # Summarize results
    res <- data.frame(
        estimator = c(
            # "True Effect", "Complete Observations",
            "Unit-wise Deletion",
            # "Bias-corrected (Infeasible)",
            # "Bias-corrected (Feasible)",
            "Mean Imputation",
            "Median Imputation"
        ),
        tau = c(
            # tau_true, fit_complete,
            fit_missing,
            # fit_missing - bias_est_complete$bias,
            # fit_missing - bias_est_missing$bias,
            fit_imp,
            fit_imp_med
        ),
        missing_var = paste0(miss_var_name, collapse = ", "),
        stringsAsFactors = FALSE
    )
    res <- res |>
        mutate(
            gamma = if_else(estimator == "Unit-wise Deletion", 0, NA_real_),
            delta = if_else(estimator == "Unit-wise Deletion", 0, NA_real_),
            bias_est = if_else(estimator == "Unit-wise Deletion", 0, NA_real_)
        )
    res <- bind_rows(
        res,
        data.frame(
            estimator = "Bias-corrected (Infeasible)",
            tau = fit_missing - bias_est_complete$bias,
            gamma = bias_est_complete$gamma,
            delta = bias_est_complete$imbalance,
            bias_est = bias_est_complete$bias,
            missing_var = paste0(miss_var_name, collapse = ", "),
            stringsAsFactors = FALSE
        ),
        data.frame(
            estimator = "Bias-corrected (Feasible)",
            tau = fit_missing - bias_est_missing$bias,
            gamma = bias_est_missing$gamma,
            delta = bias_est_missing$imbalance,
            bias_est = bias_est_missing$bias,
            missing_var = paste0(miss_var_name, collapse = ", "),
            stringsAsFactors = FALSE
        )
    )
    return(res)
}

#' Function to generate combinations of k elements
#' 
#' @param input_vector A character vector of input elements
#' @param k An integer of the number of elements to choose
generate_combinations <- function(input_vector, k) {
    # Create a list of combinations
    combinations <- combn(input_vector, k, simplify = FALSE)
    return(combinations)
}
#' Function to prompt the user to proceed or not
proceed_or_not <- function() {
  # Display a prompt to the user with a detailed message
  response <- readline(prompt = "Warning: This process might take several minutes to complete. Are you sure you want to continue? (yes/no): ")

  # Convert the response to lowercase and trim spaces to handle variations
  response <- tolower(trimws(response))

  # Check if the user wants to proceed
  if (response %in% c("yes", "y")) {
    # Code to run if the user chooses to proceed
    message("Proceeding with the operation...")
    # Your task that may take some time goes here
  } else {
    # Code to run if the user chooses not to proceed
    message("Operation aborted by the user.")
  }
}

#' Function to replace each Xn with corresponding country
replace_labels <- function(x, labels) {
  sapply(str_split(x, ",\\s*"), function(v) paste(labels[v], collapse = ", "))
}
```

Now, we run a single simulation with two units with missing outcomes.

```{r run_single_simulation}
# Run a single simulation with 2 units with missing outcomes
sim_res <- run_ife_simulation(synth_ife_data, miss_index = 1:50, k = 2)
# Check: print the results
sim_res |>
  head(8) |>
  knitr::kable(digits = 3)
```
